\section{Introduction}

Since digital communication became common, proliferation and spreading of \textit{Children Sexual Abuse Material (CSAM)} became a growing problem that required the intervention of both states and service providers. 

In response to this phenomenon, on 11 May 2022, the European Union Commission adopted a legislative proposal to fight the spreading of CSAM and online children sexual soticitation (grooming)\cite{eu2023chatcontrol}. This proposed regulation is also known as Chatcontrol.

The aim of this legislation is to detect these crimes by applying technologies based on ML classifiers for images and text on intercepted communications on targeted digital platforms. 

However, such proposal raised debates and subsequent alarms on the potential effects that this proposal might cause on the privacy of the European citizens. In particular, concern was caused by the proposed scanning methodology that would endanger the use of E2EE (End-to-End Encription) \cite{effects}.

This work aims at exploring not only the legislative aspects of this proposal relating to the privacy rights in the EU by studying the proportionality of the legislation, but also to explore the technical feasibility, efficiency, and efficacy of the proposed infrastructure in solving the presented problem.

In Section \ref{s:related} this discussion firstly present some relevant technical aspects relating to content-scanning technologies and algorithms. Successively, the topic is shifted to introduce parts of the European lesgislative framework concerning data and communication privacy. Thereafter, the proposal object of this work is presented.

Moreover, In Section \ref{s:meth} the methodology used for gathering the data is discussed.

In Section \ref{s:find} will show considerations on how the proposed system could raise an alarming number of false-posities, how it would perfrom in an adversarial context, and how detection could be avoided.

Furthermore, an exploration of some judicial cases from the \textit{European Court of Human Rights (ECHR)} is presented to understand how privacy rights, and national security and other social matters are balanced.

Successively, in Section \ref{s:disc} a discussion on the legislation proportionality is given based on the results of the evaluation of the above cited legislative framework, court rulings, and technical assessments on efficacy and efficiency, with also an opinion on a possible outcome for E2EE technology.

Finally, in Section \ref{s:concl} the conclusion is presented calling for further research in balancing child safety and privacy rights, and on less intrusive technologies for crime detection in digital communications.